---
title: "ggtree-steps"
output: html_document
---

##Using ggtree to redo figures for Darrin's paper.

```{r}
setwd("/Users/michael_alfaro/Dropbox/git/mbuna-mpe")
library(ape)

```

```
/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/manuscript-final-formatting/Manuscript Version for Submission/
```
On Dec 6 2016 I added this file to the git repository at `/Users/michael_alfaro/Dropbox/git/acanthomorphs/acanthomorph_figure_april_2015.ipynb` locally.
'
##example of using taxize

Want to count the number of families in the tree

```{r}

library(ape)
library(taxize)

tt <- read.tree("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/trees/75p/ExaBayes_ConsensusExtendedMajorityRuleNewick.Acanthomorph-75p-STDPART-1.5M-Burn25-FINAL.tre")

species.list <- tt$tip.label
temp <- gnr_resolve(names = species.list)
head(temp)
fams <- tax_name(query = species.list, get = "family", db = "ncbi")
unique(fams)


beasttree <- read.beast("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/trees/75p/ExaBayes_ConsensusExtendedMajorityRuleNewick.Acanthomorph-75p-STDPART-1.5M-Burn25-FINAL.tre")
taxonomy <- read.csv("~/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tree-taxonomy")
castax <- read.csv("~/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tree-taxonomy-cas-txt")
etoltax <- read.csv("~/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tree-taxonomy-etol-txt")

reducedetol <- select(etoltax,order, family, full.species ) 

length(unique(reducedetol$family))
length(unique(reducedetol$order))
##get rid of index column
castax <- castax[,-1]
head(castax)
p <- p %<+% castax
p <- p %<+% reducedetol
head(p$data)
# Read in trees
Get PAML label tree, and append bootstrap and bayes values to it
```{r}

library(ggtree)
library("ggplot2")  
library("colorspace")
library("gridExtra")
library(dplyr)

setwd("~/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/")
##The tree
#beast <- read.beast("final-output.tre")
beast <- read.beast("test4.tre")
#beast <-  groupClade(beast, node = 195)

#p <- ggtree(beast, right = TRUE, timescale = TRUE, aes(color=group, linetype=group)) 
p <- ggtree(beast, right = TRUE, timescale = TRUE) 
p
##tree taxonomy
taxonomy <- read.csv("~/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tree-taxonomy")
castax <- read.csv("~/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tree-taxonomy-cas-txt")
  etoltax <- read.csv("~/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tree-taxonomy-etol-txt")


reducedetol <- select(etoltax, tip.name, infraclass, megacohort, supercohort, subcohort, infracohort, section, subsection, division, subdivision, series, order, family, full.species ) 
##get rid of index column
castax <- castax[,-1]
head(castax)
p <- p %<+% castax
p <- p %<+% reducedetol
head(p$data)

##write the tips and the taxonomy to a file
reduced_data <- p$data
temp <- select( reduced_data, label, family.x, genus, full.species)
temp <- na.omit(temp)

write.table(file="sampled_species.csv", temp, row.names = F, quote = F, sep = "\t")



p <- p + geom_text(aes(label=genus), size = 1, fontface = 3)

p <- p + geom_text(aes(label=family.x), hjust = 2, color = "purple", size = 2) + geom_text(aes(label=order.x), hjust = 4, color = "red", size = 3) + geom_text(aes(label=order.y), hjust = 4, vjust = 2, color = "chocolate", size = 3) + geom_text(aes(label=series), hjust = 8, color = "chocolate2", size = 3)  + geom_text(aes(label=subdivision), hjust = 8, vjust = 4, color = "chocolate4", size = 3)
p
p$data$x <- p$data$x - max(p$data$x)
p <- p +  scale_x_continuous(breaks=seq(0,-150, -20), minor_breaks=seq(0, -150, -5)) 
p <- p + theme_tree2() +theme(panel.grid.major   = element_line(color="black", size=.2), 
                      panel.grid.minor   = element_line(color="grey", size=.1),
                      panel.grid.major.y = element_blank(),
                      panel.grid.minor.y = element_blank())  

p <- p +  geom_nodepoint(aes(color=PP, shape=ifelse(bootstrap < 70, "18", "16")), size = 6) + scale_color_gradient(low="red", high="black")  + geom_text(aes(label=pamllabel), col = "white", size = 2)
img_info <- read.csv("~/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tip-image-paths.csv", header = F, as.is = T)
head(img_info)
imgshort <- img_info[1:2,]
imgshort
#p <-  annotation_image(p, imgshort, width = 0.1)
p <-  annotation_image(p, img_info, width = 0.1)
p <- annotation_clade(p, node=106, "Acanthomorpha", offset = 40, offset.text=3, size = 3) %>% 
annotation_clade(node=195, "acanthuroids+", offset = 20, offset.text=3, size = 2) %>% 
annotation_clade(node=184, "Percomorpha", offset = 30, offset.text=3) %>% 
annotation_clade(node=107, "Paracanthopterygii", offset = 30, offset.text=3) %>%
annotation_clade(node=158, "Ovalentaria", offset = 30, offset.text=3) %>%
annotation_clade(node=167, "Carangimorpha", offset = 30, offset.text=3) %>%
annotation_clade(node=139, "Scombrimorpha", offset = 30, offset.text=3) %>%
annotation_clade(node=125, "Gobiaria", offset = 30, offset.text=3) 

p
#saving options
pdf("input-fish-tax.pdf", height=15)
p
dev.off()
p


```

##making a plot of the ages by node

read in the summary of the mcmctree combined output

```
library(ggplot2)
library(tidyr)
agesum <- read.csv("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/paml-age-summaries/4-part-ages.csv", header = T,  as.is = T)
prior_agesum <- read.csv("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/paml-age-summaries/PRIOR-4-part-ages.csv", header = T,  as.is = T)

head(agesum)
head(prior_agesum)

names(agesum)
nodes <- names(agesum[-1])
stats <- agesum[,1]
agelong <- t(agesum)
agelong <- agelong[-1,]
colnames(agelong) <- stats
head(agelong)
agelong <- agelong[,-6]
head(agelong)
dat[, c(1,23,4)] <- sapply(dat[, c(3,6:15,37)], as.numeric)


#convert to long format
agelong <- setNames(data.frame(t(agelong[,-1])), agelong[,1])
rownames(agelong) <- agelong[,1]
agelong <- agelong[,-1]
names(agelong) <- names(agesum[-1])
head(agelong)


names(agesum)

agelong <- t(agesum)

setNames(data.frame(t(mydf[,-1])), mydf[,1])
```

#p <- p +  geom_nodepoint(aes(color=PP, shape=ifelse(bootstrap < 70, "18", "16")), size = 6)  + geom_text(aes(label=pamllabel), col = "white", size = 2)
#p <- p+theme(legend.position="right")


#images need to come after 

p + geom_text(aes(label=node), hjust = 1, color = "red") # to get ggplot node numbers

#labelling nodes of interest
# 107 = paracanthopterygians
# 158 = Ovalentaria
# 167 = Carangimorpha
# 195 Acanthuroids +
# 139 Scombrimorphs
# 125 Gobiiformes

temp <- p
p <- temp




p <- groupClade(p, node=195)





p <- groupClade(p, node=213)
ggtree(p, aes(color=group, linetype=group))

#follow this example
tree <- groupClade(tree, node=c(21, 17))
ggtree(tree, aes(color=group, linetype=group))

```


theme_tree2() +
    geom_text(aes(x=max(x), label=label), size=1, color=cols, hjust=-.3) +
        scale_x_continuous(breaks=c(1992, 1995, 1997, 2000, 2002, 2005), minor_breaks=seq(1992, 2005, 1)) +
            geom_segment(aes(xend=max(x)+.20, yend=y), linetype="dotted", size=.1, color=cols) +
                theme(panel.grid.major   = element_line(color="black", size=.2),
                      panel.grid.minor   = element_line(color="grey", size=.2),
                      panel.grid.major.y = element_blank(),
                      panel.grid.minor.y = element_blank())



p <- p %<+% taxonomy + geom_text(aes(label=Family, size=1)) 

head(p$data)

geom_nodepoint(,color="black", alpha=1, size=6)   + geom_text(aes(label=pamllabel), size=2, color = "white")


ggtree(beast) + geom_text(aes(label=PP), hjust=1.25) + theme_tree2()  + geom_text(aes(label=pamllabel), col = "red", hjust=-0.1) + geom_text(aes(label=label), size=3, color="purple", hjust=-0.3)







ggtree(beast) + geom_text(aes(label=ifelse(bootstrap < 90, bootstrap, NA), col=bootstrap/100), vjust=1, hjust=1.25) + geom_text(aes(label=ifelse(PP< 0.95, PP, NA), col=PP), vjust=-0.5, hjust=1) + geom_text(aes(label=pamllabel), col = "black", hjust=-0.1) + scale_color_gradient(low="red", high="black")



ggtree(beast, aes(color=PP)) + scale_color_continuous(low="green", high="red") + geom_nodepoint(aes(size=bootstrap), alpha=1/4) + scale_color_continuous(low="yellow", high="blue") + theme(legend.position="right")



##from JC

ggtree(beast) + geom_text(aes(label=bootstrap, col=bootstrap/100), vjust=1, hjust=1) + geom_text(aes(label=PP, col=PP), vjust=-1, hjust=1) + geom_text(aes(label=pamllabel), col = "black", hjust=-0.1) + scale_color_gradient(low="red", high="black")


# 
# bayes <- read.beast("output-bayes.tre")
# boot <- read.beast("output-boot.tre")
# 
# 
# p <- ggtree(bayes, right = TRUE, timescale = TRUE)  
# p$data$x <- p$data$x - max(p$data$x)
# 
# p + geom_text(aes(label=PP))
# 
# p <- ggtree(boot, right = TRUE)  
# p$data$x <- p$data$x - max(p$data$x)
# 
# p + geom_text(aes(label=bootstrap))
# 
# 
# 
# 
# 
# 
# mergedbeast <- merge_tree(beast,bayes)
# mergedbeast <- merge_tree(mergedbeast, boot)
# 
# p <- ggtree(mergedbeast, right = TRUE, timescale = TRUE)  
# 
# head(p$data)
# test_tree <- merge_tree(beast,bayes) %>% merge_tree(boot)

```

append tree taxonomy and image information

```{r}
# 
# ##tree taxonomy
# taxonomy <- read.csv("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tree-taxonomy")
# ##get rid of index column
#  taxonomy <- taxonomy[,-1]
# head(taxonomy)
# 
# 
# ###images
# img_info <- read.csv("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tip-image-paths.csv", header = F, as.is = T)
# head(img_info)
# 
# imgshort <- img_info[1,]
# imgshort
# #ggtree(beast, right = TRUE) + geom_text(aes(label = node, color = isTip)) + scale_color_manual(values = c("black", NA))
# ggtree(beast, right = TRUE) # + geom_text(aes(label = node, color = isTip)) + scale_color_manual(values = c("black", NA))
# # ggtree(beast, right = TRUE) + geom_text(aes(label = node, color = isTip)) + scale_color_manual(values = c("black", NA))
# p$data$x <- p$data$x - max(p$data$x)
# 
# p + geom_text(aes(label=bootstrap))
# p + geom_text(subset=.(bootstrap>69), aes(label=bootstrap))
# 
# p + geom_text(subset=.(PP>69), aes(label=PP))
# 
# 
# #p + geom_nodepoint(aes(size=PP), alpha=0.25) + scale_color_continuous(low="green", high="red")  
# 
# p <- p +  geom_nodepoint(,color="black", alpha=1, size=6)   + geom_text(aes(label=pamllabel), size=2, color = "white", fontface = 3) + theme_tree2(legend.position="right")
# p

```



p <- p +  geom_nodepoint(subset=.(bootstrap>50),color="black", alpha=1, size=6) + scale_color_gradient(limits=c(50, 100), low="red", high="black")  + geom_text(aes(label=pamllabel), size=2, color = "white") + theme_tree2(legend.position="right")
p <- p %<+% taxonomy + geom_text(aes(color=Order, label=Family, size=1)) 

p <- p +  scale_x_continuous(breaks=seq(0, 150, 10), minor_breaks=seq(-150, -0, -5)) 
p <- p + theme(panel.grid.major   = element_line(color="black", size=.2),
                      panel.grid.minor   = element_line(color="grey", size=.2),
                      panel.grid.major.y = element_blank(),
                      panel.grid.minor.y = element_blank())


##get bootstrap and PP values on to tree



#p+ xlim(0,300)
#p <- annotation_image(p, img_info, width = 0.2)
#p <- geom_text(aes(x=node, label = "pamllabel", color = "red"))
 

ggtree(ml, branch.length="dN_vs_dS") + 
    geom_text(aes(x=branch, label=dN), 
              size=3, vjust=-0.5, color="red") +
                  geom_text(aes(x=branch, label=dS),
                            size=3, vjust=1.2, color="darkgreen")






###this function gets the daughter tips from a node

para <- get.offspring.tip(beast@phylo, 120)
gobioids <- get.offspring.tip(beast@phylo, 139)
tumpts <- get.offspring.tip(beast@phylo, 154)



```{r eval=FALSE}
# cls <- list(c1=c("A", "B", "C", "D", "E"),
#             c2=c("F", "G", "H"),
#             c3=c("L", "K", "I", "J"),
#             c4="M")
# 
# tree <- groupOTU(tree, cls)
# 
# ggtree(tree, aes(color=group, linetype=group)) + geom_text(aes(label=label),  hjust=-.25) +
#      scale_color_manual(values=c("black", rainbow_hcl(4))) + theme(legend.position="right")
# 
# 
# 
# 
# beast <- read.beast("labeled-node-tree.tre")
# #ladderize(beast)
# plot(beast, right = TRUE, annotation="pamllabel", annotation.color="red") + theme_tree() 
# 
# ggtree(beast, color="steelblue",right = TRUE, size=0.5, linetype="dotted")
# ggtree(beast, setbranchlength = "none")
# 
# grid.arrange(ggtree(beast) + ggtitle("(Phylogram) rectangular layout"),
#              ggtree(beast, branch.length='none') + ggtitle("(Cladogram) rectangular layout"),
#          ggtree(beast, layout="slanted") + ggtitle("(Phylogram) slanted layout"),
#              ggtree(beast, layout="slanted", branch.length='none') + ggtitle("(Cladogram) slanted layout"),
#          ggtree(beast, layout="circular") + ggtitle("(Phylogram) circular layout"),
#              ggtree(beast, layout="circular", branch.length="none") + ggtitle("(Cladogram) circular layout"),
#          ggtree(beast, layout="unrooted") + ggtitle("unrooted layout"),
#          ncol=2)
# 
# ggtree(beast) %>% add_legend()
# ggtree(beast) + theme_tree2()
# 
# ggtree(beast)+geom_point(aes(shape=isTip, color=isTip), size=3)
# 
# p <- ggtree(beast) + geom_nodepoint(color="#b5e521", alpha=1/4, size=10)
# p + geom_tippoint(color="#FDAC4F", shape=8, size=3)
# p + geom_text(aes(label=label), size=3, color="purple", hjust=-0.3)
# ggtree(beast, layout="circular") + geom_text(aes(label=label, angle=angle), size=3, color="purple", vjust=-0.3)
# 
# p + geom_text(aes(x=branch, label=label), size=3, color="purple", vjust=-0.3)
# 
# grid.arrange(
#     ggtree(beast, color="red") + theme_tree2("steelblue"),
#     ggtree(beast, color="white") + theme_tree("black"),
#     ncol=2)
# p %<% rtree(50)
# 
# gzoom(beast, grep("canthigaster", beast@phylo$tip.label)) #this would be handy for large trees--need a way to generate list of tips from a node number
# 
# ggtree(beast, aes(color=branch.length)) +
#     scale_color_continuous(low="green", high="red") +
#         theme(legend.position="bottom")
# 
# ggtree(beast) + geom_text(aes(label=node)) + geom_tiplab()
# 
# p <- ggtree(beast) + geom_point(aes(shape=isTip, color=isTip), size=3) 
# annotation_clade(p, node=120, "paracanth", offset.text=50)
# 
# p <- ggtree(beast) 
# annotation_clade2(p, "lampris_guttatus", "aphredoderus_sayanus", "Clade X", offset.text=3) %>%
#     annotation_clade2("parachirus_xenicus", "chloroscombrus_orqueta", "Clade Y", bar.size=4, font.size=8, offset=5, offset.text=4, color="steelblue")
# 
# ggtree(beast) %>% hilight(node=120, fill="steelblue", alpha=.6) %>%
#     hilight(node=200, fill="darkgreen", alpha=.6)
# 
# ggtree(beast, layout="fan") %>% hilight(node=120, fill="steelblue", alpha=.6) %>%
#      hilight(node=200, fill="darkgreen", alpha=.6)
# 
# cp <- ggtree(beast) %>% collapse(node=21)
# cp + geom_point(subset=.(node == 21), size=5, shape=23, fill="steelblue")
# ##
# 
# bb <- groupOTU(beast, focus=c("lampris_guttatus", "zu_elongatus"))
# bb <- ggtree(bb, aes(color=group))
# 
# bb + geom_text(aes(label=label), size=3, color="purple", hjust=-0.3)
# 
# 
# bb <- groupClade(beast, node=c(200, 150))
# ggtree(bb, aes(color=group, linetype=group))
# 
# ggtree(bb, aes(color=group, linetype=group)) +
#     geom_text(aes(label=label), subset=.(group==2), hjust = -.5) +
#         geom_text(aes(label=label), subset=.(group==1), hjust = -.5, color="black")
# 
# pp <- ggtree(beast) %>% phylopic("cb66c697-7431-4368-b134-eb9019e2ae10", color="steelblue", alpha = .3)
# print(pp)
# 
# pp %>% phylopic("cb66c697-7431-4368-b134-eb9019e2ae10", color="#86B875", alpha=.8, node=4) %>%
# +      phylopic("8d92b454-3131-4bbd-ac9c-e1df14c2fc5a", color="darkcyan", alpha=.8, node=17, width=.2)
# 
# 
# ##annotating node with taxonomy
# ##read in data from python notebook 12_calibrations_PAML_Sept_2015
# 
# df <- read.csv("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tree-taxonomy")
# 
# ##get rid of index column
#  df <- df[,-1]
# head(df)
# 
# 
# ## the  %<+% operator attaches a data frame to the tips of a tree view
# 
# 
# ##the example
# nwk <- system.file("extdata", "sample.nwk", package="ggtree")
# tree <- read.tree(nwk)
# p <- ggtree(tree)
# 
# dd <- data.frame(taxa  = LETTERS[1:13], 
#              place = c(rep("GZ", 5), rep("HK", 3), rep("CZ", 4), NA),
#                  value = round(abs(rnorm(13, mean=70, sd=10)), digits=1))
# ## you don't need to order the data
# ## data was reshuffled just for demonstration
# dd <- dd[sample(1:13, 13), ]
# row.names(dd) <- NULL
# 
# p <- p %<+% dd + geom_text(aes(color=place, label=label), hjust=-0.5) + 
#        geom_tippoint(aes(size=value, shape=place, color=place), alpha=0.25)
# p+theme(legend.position="right")
# 
# 
# # p  + geom_tiplab() + ggplot2::xlim(0, 0.06) + ggplot2::scale_x_continuous
# # annotation_image(p, imgshort,  width = 0.05, linetype=NULL, offset=.1)
# #annotation_image(p, imgshort,  width = 0.05, linetype=NULL, offset=.1)  + ggplot2::xlim(0, 1)
# 
# # This is because the units are in two different spaces (data & pixel).
# # 
# # you can use xlim to get more space for text.
# # 
# # try:
# # 
# # ggtree(tree) + geom_tiplab() + ggplot2::xlim(0, 0.06)
# 
# 
# # 
# # img_info <- read.csv("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/dating/Friedman_12_calibrations/PAML_run_files/4_partitions/annotated_figtree_work/full-tree-work/tip-image-paths.csv", header = F, as.is = T)
# # head(img_info)
# # 
# # 
# # p <- ggtree(beast, time_scale=T,right = T )
# # setwd("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/manuscript/pnas_tex/ETE_work/converted_pngs/")
# # dfshort <- df[1,]
# # 
# # annotation_image(p, img_info,  width = 0.2, linetype=NULL, offset=.3)
# # 
# # 
# # p <- p %<+% df + geom_text(aes(color=Family, label=Genus), hjust=-0.5)
# # p
# # 
# 
# 

```

###get the tree loaded
I have a figtree file. 
make a file with the taxonomy for each tip

```{r}
# library(ape)
# beast <- read.beast("labeled-node-tree.tre")
# exa <- read.tree("/Users/michael_alfaro/Dropbox/malfaro-acanthomorph/trees/75p/ExaBayes_ConsensusExtendedMajorityRuleNewick.Acanthomorph-75p-STDPART-1.5M-Burn25-FINAL.tre")
# #ladderize(beast)
# 
# 
# # 
# #remove outgroups here
# exa <- drop.tip(exa, c("alepisaurus_ferox", "ceratoscopelus_warmingii" ))
# 
# 
# layout(matrix(1:2, 1, ))
# plot(beast@phylo, type = "phylogram", right = T,  use.edge.length = F, show.node.label = F, cex = 0.5)
# nodelabels(frame = "none", cex = 0.5)
# plot(exa, type = "phylogram", right = T, use.edge.length = F, show.node.label = F, cex = 0.5)
# nodelabels(frame = "none", cex = 0.5)
# 
#  exa <- reroot(exa, 225)
# # 
# # layout(matrix(1:2, 1, ))
# # plot(beast@phylo, type = "phylogram", use.edge.length = F, show.node.label = F, cex = 0.5)
# # nodelabels(frame = "none", cex = 0.5)
# # plot(exa, type = "phylogram", use.edge.length = F, show.node.label = F, cex = 0.5)
# # nodelabels(frame = "none", cex = 0.5)
# # 
# # 
# # exa <- reroot(exa, 224)
# # 
#  mt <- merge_tree(beast, exa)
# # 
# # beastplot(beast, right = TRUE, annotation="HPD", annotation.color="red") + theme_tree() 
# 
# #read in a tree with posterior probabilities

```

